---
title: "Hampton Childcare — Analysis"
date: "`r Sys.Date()`"
format: gfm
---

## Overview

Applies project-specific filters and computes the headline metrics for the
Hampton childcare access analysis, in response to findings from the 2024
Hampton Poverty Study. The analysis focuses on under-5 facilities — those
whose youngest served age is below 60 months.

Four metrics are computed:

1. **Subsidy acceptance** — share of under-5 facilities accepting the Virginia
   Child Care Subsidy program
2. **Infant service** — share of under-5 facilities serving infants (age min < 12 months)
3. **Nontraditional hours** — share operating between 8 pm and 5 am (or spanning midnight)
4. **Weekend operation** — share open on Saturday or Sunday

Nontraditional hours and weekend counts are expressed as ranges because CCAV
and DSS hours data sometimes disagrees. The **low** estimate requires both
sources to indicate nontraditional/weekend operation; the **high** estimate
fires if either source does. Phone-confirmed values are treated as fixed.

**Input:**
- `Data/merge/providers_merged_YYYY-MM-DD.csv`
- `Data/merge/providers_hours_merged_YYYY-MM-DD.csv`
- `Data/merge/manual_overrides_hours.csv`
- `Data/merge/manual_overrides_subsidy.csv`

**Output:** `Data/analysis/providers_analysis_YYYY-MM-DD.csv`

| Limitation | Detail |
|---|---|
| CCAV-only facilities have unknown subsidy status | 37% figure is based only on facilities with a DSS record. True share may differ. |
| Missing hours data | Facilities with no hours in either source are excluded from nontraditional/weekend counts. Best available data. |
| ECSE / special needs | Derived from CCAV `program_type` free text only. DSS carries no equivalent field. Likely an undercount. |
| YMCA manual exclusion | age_min = 57 months technically passes the < 60 cutoff but the program is not an early-childhood provider. |

---

```{r setup, include=FALSE}
library(here)
library(readr)
library(dplyr)
library(stringr)
library(purrr)
library(knitr)

source(here::here("config.R"))
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

today <- format(Sys.Date(), "%Y-%m-%d")
```

## Helper Functions

```{r helpers, include=FALSE}
locate_latest <- function(dir, pattern) {
  f <- list.files(dir, pattern = pattern, full.names = TRUE)
  if (!length(f)) stop("No file matching '", pattern, "' in ", dir)
  f[which.max(file.mtime(f))]
}

# Convert HH:MM string to total minutes since midnight
to_min <- function(t) {
  t <- str_extract(t, "^[0-9]{2}:[0-9]{2}")
  if_else(is.na(t), NA_real_,
    as.numeric(str_sub(t, 1, 2)) * 60 + as.numeric(str_sub(t, 4, 5)))
}

# A facility day is nontraditional if it closes at or after 8 pm, opens at or
# before 5 am, or spans midnight (close < open). Threshold chosen to match the
# 2024 Hampton Poverty Study caregiver survey definition (8 pm–5 am window).
is_nontrad <- function(open_t, close_t) {
  op <- to_min(open_t)
  cl <- to_min(close_t)
  (!is.na(cl) & cl >= 20 * 60) |
  (!is.na(op) & op <=  5 * 60) |
  (!is.na(op) & !is.na(cl) & cl < op)
}

# Compute nontraditional and weekend flags per facility from the long hours table.
# Returns one row per name_key with four boolean columns:
#   nontrad_low  / nontrad_high — conservative / liberal nontraditional estimate
#   weekend_low  / weekend_high — conservative / liberal weekend estimate
compute_facility_hours <- function(h) {
  h |>
    mutate(
      # Confirmed rows are treated as authoritative regardless of discrepancy
      nt_confirmed   = hours_confirmed & coalesce(is_open, FALSE) & is_nontrad(open_time, close_time),
      # Coalesced rows: unconfirmed, no discrepancy — use merged time
      nt_coalesced   = !hours_confirmed & !coalesce(time_discrepancy, FALSE) &
                         coalesce(is_open, FALSE) & is_nontrad(open_time, close_time),
      nt_ccav        = is_nontrad(open_time_ccav, close_time_ccav) & coalesce(is_open_ccav, FALSE),
      nt_dss         = is_nontrad(open_time_dss,  close_time_dss)  & coalesce(is_open_dss,  FALSE),
      # Low: both sources agree it's nontraditional; High: either source says so
      nt_disc_both   = !hours_confirmed & coalesce(time_discrepancy, FALSE) & nt_ccav & nt_dss,
      nt_disc_either = !hours_confirmed & coalesce(time_discrepancy, FALSE) & (nt_ccav | nt_dss),
      is_weekend     = day %in% c("Saturday", "Sunday"),
      wknd_open_conf = is_weekend & hours_confirmed & coalesce(is_open, FALSE),
      wknd_open_coal = is_weekend & !hours_confirmed &
                         !coalesce(time_discrepancy, FALSE) & coalesce(is_open, FALSE),
      wknd_ccav      = is_weekend & !hours_confirmed & coalesce(is_open_ccav, FALSE),
      wknd_dss       = is_weekend & !hours_confirmed & coalesce(is_open_dss,  FALSE)
    ) |>
    group_by(name_key) |>
    summarise(
      nontrad_low  = any(nt_confirmed | nt_coalesced | nt_disc_both,   na.rm = TRUE),
      nontrad_high = any(nt_confirmed | nt_coalesced | nt_disc_either, na.rm = TRUE),
      weekend_base = any(wknd_open_conf | wknd_open_coal,             na.rm = TRUE),
      weekend_ccav = any(wknd_ccav,                                   na.rm = TRUE),
      weekend_dss  = any(wknd_dss,                                    na.rm = TRUE),
      .groups = "drop"
    ) |>
    mutate(
      weekend_low  = weekend_base | (weekend_ccav & weekend_dss),
      weekend_high = weekend_base | weekend_ccav | weekend_dss
    )
}

pct     <- function(n, d) sprintf("%.0f%%", 100 * n / d)
pct_rng <- function(lo, hi, d) sprintf("%.0f%%–%.0f%%", 100 * lo / d, 100 * hi / d)
```

## Load Data

```{r load-data}
merged_path       <- locate_latest(here("Data", "03_merge"), "^providers_merged_.*\\.csv$")
hours_merged_path <- locate_latest(here("Data", "03_merge"), "^providers_hours_merged_.*\\.csv$")

cat(sprintf("Merged:       %s\nHours (long): %s\n",
  basename(merged_path), basename(hours_merged_path)))

merged       <- read_csv(merged_path,       show_col_types = FALSE)
hours_merged <- read_csv(hours_merged_path, show_col_types = FALSE)

hours_ov_path     <- here("Data", "overrides", "manual_overrides_hours.csv")
subsidy_ov_path   <- here("Data", "overrides", "manual_overrides_subsidy.csv")
hours_overrides   <- if (file.exists(hours_ov_path))   read_csv(hours_ov_path,   show_col_types = FALSE) else tibble()
subsidy_overrides <- if (file.exists(subsidy_ov_path)) read_csv(subsidy_ov_path, show_col_types = FALSE) else tibble()
```

## Apply Subsidy Overrides

```{r subsidy-overrides}
if (nrow(subsidy_overrides) > 0) {
  bad_sub_ov <- subsidy_overrides |>
    anti_join(merged |> distinct(name_key), by = "name_key")
  if (nrow(bad_sub_ov) > 0) {
    warning("Subsidy override name_key(s) not found in merged: ",
            paste(bad_sub_ov$name_key, collapse = ", "))
  }
  merged <- merged |>
    left_join(
      subsidy_overrides |>
        filter(!is.na(accepts_subsidy)) |>
        select(name_key, .sub_ov = accepts_subsidy),
      by = "name_key"
    ) |>
    mutate(
      subsidy_assumed = !is.na(.sub_ov),
      accepts_subsidy = if_else(subsidy_assumed, .sub_ov, accepts_subsidy)
    ) |>
    select(-.sub_ov)
  cat(sprintf("Applied %d subsidy overrides.\n", sum(merged$subsidy_assumed)))
} else {
  merged <- merged |> mutate(subsidy_assumed = FALSE)
  cat("No subsidy overrides.\n")
}
```

## Apply Exclusions

```{r exclusions}
# The YMCA (young mens christian) has age_min = 57 months — just under the 60
# month cutoff — but is not an early-childhood childcare provider. Manually
# excluded to avoid inflating the under-5 count.
manual_exclude        <- c("young mens christian")
manual_exclude_reason <- "age_min = 57 months (4yr 9mo); known non-early-childhood program"

n_before   <- nrow(merged)
analysis   <- merged |>
  filter(
    !str_detect(name_key, regex(paste(manual_exclude, collapse = "|"), ignore_case = TRUE)),
    !is.na(age_min_months),
    age_min_months < 60
  )

n_excl_age <- sum(!is.na(merged$age_min_months) & merged$age_min_months >= 60)
n_excl_na  <- sum(is.na(merged$age_min_months))
n_excl_man <- sum(str_detect(merged$name_key,
                   regex(paste(manual_exclude, collapse = "|"), ignore_case = TRUE)))
na_names   <- merged |> filter(is.na(age_min_months)) |> pull(name) |> paste(collapse = "; ")
```

## Compute Hours Metrics

```{r compute-hours}
u5  <- analysis |> filter(serves_under_5 == TRUE)
u5s <- u5       |> filter(accepts_subsidy == TRUE)

hours_analysis <- hours_merged |> filter(name_key %in% analysis$name_key)
hours_u5       <- hours_merged |> filter(name_key %in% u5$name_key)
hours_u5s      <- hours_merged |> filter(name_key %in% u5s$name_key)

fac_hours     <- compute_facility_hours(hours_analysis)
fac_hours_u5  <- compute_facility_hours(hours_u5)
fac_hours_u5s <- compute_facility_hours(hours_u5s)
```

## Write Output

```{r write-output}
analysis_out <- analysis |>
  left_join(
    fac_hours |> select(name_key, nontrad_low, nontrad_high, weekend_low, weekend_high),
    by = "name_key"
  )

output_path <- here("Data", "04_analyze", paste0("providers_analysis_", today, ".csv"))
dir.create(here("Data", "04_analyze"), showWarnings = FALSE, recursive = TRUE)
write_csv(analysis_out, output_path)
prune_dated(here("Data", "04_analyze"), "^providers_analysis_[0-9]{4}-[0-9]{2}-[0-9]{2}\\.csv$")
cat(sprintf("Wrote %d rows → %s\n", nrow(analysis_out), basename(output_path)))
```

## Main Findings

These are the headline metrics that motivated this analysis. All figures are
for under-5 facilities only (age_min < 60 months, YMCA excluded).

```{r main-findings}
n_total   <- nrow(analysis)
n_u5      <- nrow(u5)
n_u5s     <- nrow(u5s)

n_u5_subsidy_true    <- sum(u5$accepts_subsidy == TRUE,  na.rm = TRUE)
n_u5_subsidy_false   <- sum(u5$accepts_subsidy == FALSE, na.rm = TRUE)
n_u5_subsidy_unknown <- sum(is.na(u5$accepts_subsidy))
n_u5_infants         <- sum(u5$serves_infants, na.rm = TRUE)
n_u5s_infants        <- sum(u5s$serves_infants, na.rm = TRUE)

nontrad_low   <- sum(fac_hours$nontrad_low)
nontrad_high  <- sum(fac_hours$nontrad_high)
u5_nt_low     <- sum(fac_hours_u5$nontrad_low)
u5_nt_high    <- sum(fac_hours_u5$nontrad_high)
u5s_nt_low    <- sum(fac_hours_u5s$nontrad_low)
u5s_nt_high   <- sum(fac_hours_u5s$nontrad_high)
u5_wknd_low   <- sum(fac_hours_u5$weekend_low)
u5_wknd_high  <- sum(fac_hours_u5$weekend_high)
u5s_wknd_low  <- sum(fac_hours_u5s$weekend_low)
u5s_wknd_high <- sum(fac_hours_u5s$weekend_high)

n_hours_disc    <- hours_analysis |> filter(time_discrepancy) |> pull(name_key) |> n_distinct()
n_hours_base    <- n_distinct(hours_analysis$name_key)

kable(data.frame(
  Metric = c(
    "Under-5 facilities (analysis universe)",
    "Accept subsidy",
    "Subsidy status unknown (CCAV only)",
    "Serve infants (age_min < 12 months)",
    "Nontraditional hours (low estimate)",
    "Nontraditional hours (high estimate)",
    "Open on weekends (low estimate)",
    "Open on weekends (high estimate)"
  ),
  Count = c(
    n_u5, n_u5_subsidy_true, n_u5_subsidy_unknown, n_u5_infants,
    u5_nt_low, u5_nt_high, u5_wknd_low, u5_wknd_high
  ),
  Pct_of_under5 = c(
    "100%",
    pct(n_u5_subsidy_true,  n_u5),
    pct(n_u5_subsidy_unknown, n_u5),
    pct(n_u5_infants, n_u5),
    pct(u5_nt_low,  n_u5), pct(u5_nt_high,  n_u5),
    pct(u5_wknd_low, n_u5), pct(u5_wknd_high, n_u5)
  )
))
```

*Nontraditional hours range reflects `r n_hours_disc` facilities with per-day
CCAV/DSS time discrepancies. Phone-confirmed values are treated as fixed.*

## Summary

### Exclusion Filter

```{r summary-exclusions}
kable(data.frame(
  Step = c(
    "Total from merge",
    "Excluded: age_min ≥ 60 months",
    "Excluded: no age data",
    "Excluded: manual (YMCA)",
    "Analysis total"
  ),
  Removed = c("—", sprintf("-%d", n_excl_age), sprintf("-%d", n_excl_na),
              sprintf("-%d", n_excl_man), "—"),
  Remaining = c(n_before, n_before - n_excl_age,
                n_before - n_excl_age - n_excl_na,
                n_before - n_excl_age - n_excl_na - n_excl_man,
                n_total),
  Note = c("", "School-age only", na_names, manual_exclude_reason, "")
))
```

### Under-5 + Subsidy Cross-Tab

```{r summary-u5-subsidy}
kable(data.frame(
  Group = c("Under-5", "Under-5 + subsidy"),
  n = c(n_u5, n_u5s),
  pct_of_u5 = c("100%", pct(n_u5s, n_u5)),
  serves_infants = c(
    sprintf("%d (%s)", n_u5_infants,  pct(n_u5_infants,  n_u5)),
    sprintf("%d (%s)", n_u5s_infants, pct(n_u5s_infants, n_u5))
  ),
  nontrad_low_high = c(
    sprintf("%d–%d (%s)", u5_nt_low,  u5_nt_high,  pct_rng(u5_nt_low,  u5_nt_high,  n_u5)),
    sprintf("%d–%d (%s)", u5s_nt_low, u5s_nt_high, pct_rng(u5s_nt_low, u5s_nt_high, n_u5))
  )
))
```

## Data Quality Audit

### Missing Hours Data

```{r audit-missing-hours}
n_analysis_missing_hours <- hours_analysis |>
  group_by(name_key) |>
  summarise(has_hours = any(!is.na(open_time)), .groups = "drop") |>
  filter(!has_hours) |>
  nrow()

cat(sprintf(
  "%d of %d analysis facilities have no hours data in either source.\n",
  n_analysis_missing_hours, n_u5
))
cat("These facilities are excluded from nontraditional and weekend counts.\n")
```

### Suspicious Flag Checks

```{r audit-flags}
# Facilities with nontrad_low = TRUE but no confirmed hours — hours from sources only
source_nontrad <- analysis_out |>
  filter(nontrad_low == TRUE) |>
  left_join(hours_analysis |>
    group_by(name_key) |>
    summarise(any_confirmed = any(hours_confirmed), .groups = "drop"),
    by = "name_key"
  ) |>
  filter(!coalesce(any_confirmed, FALSE))

if (nrow(source_nontrad) > 0) {
  cat(sprintf(
    "%d facilities flagged nontraditional (low) based on unconfirmed source hours:\n",
    nrow(source_nontrad)
  ))
  kable(source_nontrad |> select(name, nontrad_low, nontrad_high, accepts_subsidy))
}
```

### Zero / Missing Value Summary

```{r audit-zeros}
kable(data.frame(
  Field = c("address", "phone", "hours (any)", "subsidy status"),
  NA_or_missing = c(
    sum(is.na(analysis_out$address)),
    sum(is.na(analysis_out$phone) | analysis_out$phone == "", na.rm = TRUE),
    n_analysis_missing_hours,
    sum(is.na(analysis_out$accepts_subsidy))
  )
))
```

### Random Spot Check

```{r audit-spot-check}
set.seed(as.integer(Sys.Date()))
kable(
  slice_sample(analysis_out, n = min(5, nrow(analysis_out))) |>
    select(name, source, age_min_months, serves_infants,
           accepts_subsidy, nontrad_low, nontrad_high),
  caption = "Random date-seeded spot check"
)
```
