---
title: "Hampton Childcare — Manual Overrides"
date: "`r Sys.Date()`"
format: gfm
---

## Overview

Stores phone-confirmed hours corrections and (currently empty) subsidy
overrides. These records are the authoritative source for any facility where
hours data from DSS or CCAV was found to be inaccurate through direct
verification.

Each `name_pattern` is a lowercase substring matched against `name_key` in the
merged provider table. A pattern may match one facility or several — the merge
module reports exactly which facilities each pattern matched. Every pattern
**must** match at least one real facility or the pipeline will stop with an
error.

**Output:**
- `Data/merge/manual_overrides_hours.csv`
- `Data/merge/manual_overrides_subsidy.csv`

| Limitation | Detail |
|---|---|
| Kimberly Glover | Voicemail left 2026-03-26; hours unconfirmed. No override rows until confirmation is received. |
| Sunday "by parent arrangement" | Annie Mae Bembury Sunday hours are indeterminate; `is_open = NA`. This is not treated as open or closed. |
| Empty subsidy overrides | No phone-confirmed subsidy status corrections exist at this time. |

---

```{r setup, include=FALSE}
library(here)
library(readr)
library(tibble)
library(dplyr)
library(stringr)
library(knitr)

source(here::here("config.R"))
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

## Phone-Confirmed Hours Overrides

Each row represents one facility × day combination confirmed by phone call.
The `confirmed_note` field records the date and substance of each call.

```{r hours-overrides}
hours_overrides <- tribble(
  ~name_pattern,             ~day,         ~open_time, ~close_time, ~is_open, ~confirmed_note,

  # annie mae bembury — Phone confirmed 2026-03-26: M-F 24hr; Sat 6am-6pm; Sun by parent arrangement
  "annie mae bembury",       "Monday",     "00:00",    "23:59",     TRUE,     "Phone confirmed 2026-03-26: M-F 24hr; Sat 6am-6pm; Sun by parent arrangement",
  "annie mae bembury",       "Tuesday",    "00:00",    "23:59",     TRUE,     "Phone confirmed 2026-03-26: M-F 24hr; Sat 6am-6pm; Sun by parent arrangement",
  "annie mae bembury",       "Wednesday",  "00:00",    "23:59",     TRUE,     "Phone confirmed 2026-03-26: M-F 24hr; Sat 6am-6pm; Sun by parent arrangement",
  "annie mae bembury",       "Thursday",   "00:00",    "23:59",     TRUE,     "Phone confirmed 2026-03-26: M-F 24hr; Sat 6am-6pm; Sun by parent arrangement",
  "annie mae bembury",       "Friday",     "00:00",    "23:59",     TRUE,     "Phone confirmed 2026-03-26: M-F 24hr; Sat 6am-6pm; Sun by parent arrangement",
  "annie mae bembury",       "Saturday",   "06:00",    "18:00",     TRUE,     "Phone confirmed 2026-03-26: M-F 24hr; Sat 6am-6pm; Sun by parent arrangement",
  # Sunday: by parent arrangement — status left as NA (not open, not closed)
  "annie mae bembury",       "Sunday",     NA,         NA,          NA,       "Phone confirmed 2026-03-26: M-F 24hr; Sat 6am-6pm; Sun by parent arrangement",

  # sylvia bennette — Phone confirmed 2026-03-26: M-F 6:30am-6pm
  "sylvia bennette",         "Monday",     "06:30",    "18:00",     TRUE,     "Phone confirmed 2026-03-26: M-F 6:30am-6pm",
  "sylvia bennette",         "Tuesday",    "06:30",    "18:00",     TRUE,     "Phone confirmed 2026-03-26: M-F 6:30am-6pm",
  "sylvia bennette",         "Wednesday",  "06:30",    "18:00",     TRUE,     "Phone confirmed 2026-03-26: M-F 6:30am-6pm",
  "sylvia bennette",         "Thursday",   "06:30",    "18:00",     TRUE,     "Phone confirmed 2026-03-26: M-F 6:30am-6pm",
  "sylvia bennette",         "Friday",     "06:30",    "18:00",     TRUE,     "Phone confirmed 2026-03-26: M-F 6:30am-6pm",
  "sylvia bennette",         "Saturday",   NA,         NA,          FALSE,    "Phone confirmed 2026-03-26: M-F 6:30am-6pm",
  "sylvia bennette",         "Sunday",     NA,         NA,          FALSE,    "Phone confirmed 2026-03-26: M-F 6:30am-6pm",

  # pure at heart childcare — Phone confirmed 2026-03-26: M-F 6am-6pm; late night/weekend not currently operating
  "pure at heart childcare", "Monday",     "06:00",    "18:00",     TRUE,     "Phone confirmed 2026-03-26: M-F 6am-6pm; late night/weekend services not currently operating",
  "pure at heart childcare", "Tuesday",    "06:00",    "18:00",     TRUE,     "Phone confirmed 2026-03-26: M-F 6am-6pm; late night/weekend services not currently operating",
  "pure at heart childcare", "Wednesday",  "06:00",    "18:00",     TRUE,     "Phone confirmed 2026-03-26: M-F 6am-6pm; late night/weekend services not currently operating",
  "pure at heart childcare", "Thursday",   "06:00",    "18:00",     TRUE,     "Phone confirmed 2026-03-26: M-F 6am-6pm; late night/weekend services not currently operating",
  "pure at heart childcare", "Friday",     "06:00",    "18:00",     TRUE,     "Phone confirmed 2026-03-26: M-F 6am-6pm; late night/weekend services not currently operating",
  "pure at heart childcare", "Saturday",   NA,         NA,          FALSE,    "Phone confirmed 2026-03-26: M-F 6am-6pm; late night/weekend services not currently operating",
  "pure at heart childcare", "Sunday",     NA,         NA,          FALSE,    "Phone confirmed 2026-03-26: M-F 6am-6pm; late night/weekend services not currently operating",

  # alstons village — Phone confirmed 2026-03-26: M-F 6am-6pm
  "alstons village",         "Monday",     "06:00",    "18:00",     TRUE,     "Phone confirmed 2026-03-26: M-F 6am-6pm",
  "alstons village",         "Tuesday",    "06:00",    "18:00",     TRUE,     "Phone confirmed 2026-03-26: M-F 6am-6pm",
  "alstons village",         "Wednesday",  "06:00",    "18:00",     TRUE,     "Phone confirmed 2026-03-26: M-F 6am-6pm",
  "alstons village",         "Thursday",   "06:00",    "18:00",     TRUE,     "Phone confirmed 2026-03-26: M-F 6am-6pm",
  "alstons village",         "Friday",     "06:00",    "18:00",     TRUE,     "Phone confirmed 2026-03-26: M-F 6am-6pm",
  "alstons village",         "Saturday",   NA,         NA,          FALSE,    "Phone confirmed 2026-03-26: M-F 6am-6pm",
  "alstons village",         "Sunday",     NA,         NA,          FALSE,    "Phone confirmed 2026-03-26: M-F 6am-6pm",

  # live laugh grow — Phone confirmed 2026-03-26: M-F 6am-6pm
  "live laugh grow",         "Monday",     "06:00",    "18:00",     TRUE,     "Phone confirmed 2026-03-26: M-F 6am-6pm",
  "live laugh grow",         "Tuesday",    "06:00",    "18:00",     TRUE,     "Phone confirmed 2026-03-26: M-F 6am-6pm",
  "live laugh grow",         "Wednesday",  "06:00",    "18:00",     TRUE,     "Phone confirmed 2026-03-26: M-F 6am-6pm",
  "live laugh grow",         "Thursday",   "06:00",    "18:00",     TRUE,     "Phone confirmed 2026-03-26: M-F 6am-6pm",
  "live laugh grow",         "Friday",     "06:00",    "18:00",     TRUE,     "Phone confirmed 2026-03-26: M-F 6am-6pm",
  "live laugh grow",         "Saturday",   NA,         NA,          FALSE,    "Phone confirmed 2026-03-26: M-F 6am-6pm",
  "live laugh grow",         "Sunday",     NA,         NA,          FALSE,    "Phone confirmed 2026-03-26: M-F 6am-6pm"
)
```

## Subsidy Overrides

Currently empty. Add entries here only when subsidy status has been confirmed
by phone or there is a documented basis for the assumption.

```{r subsidy-overrides}
subsidy_overrides <- tribble(
  ~name_pattern, ~accepts_subsidy, ~subsidy_note
  # Format: lowercase substring of name_key, TRUE/FALSE, "basis for assumption"
)
```

## Write Outputs

```{r validate-and-write}
# All name_patterns must be lowercase — prevents silent mismatch with name_key
non_lowercase <- hours_overrides |>
  distinct(name_pattern) |>
  filter(name_pattern != str_to_lower(name_pattern))

if (nrow(non_lowercase) > 0) {
  stop("Override name_pattern(s) must be lowercase:\n",
       paste(non_lowercase$name_pattern, collapse = "\n"))
}

hours_path   <- here("Data", "overrides", "manual_overrides_hours.csv")
subsidy_path <- here("Data", "overrides", "manual_overrides_subsidy.csv")

write_csv(hours_overrides,   hours_path)
write_csv(subsidy_overrides, subsidy_path)

cat(sprintf("Wrote %d rows → %s\n", nrow(hours_overrides),   basename(hours_path)))
cat(sprintf("Wrote %d rows → %s\n", nrow(subsidy_overrides), basename(subsidy_path)))
```

## Main Findings

```{r main-findings}
kable(
  hours_overrides |>
    group_by(name_pattern) |>
    summarise(
      days_confirmed = sum(!is.na(is_open) & is_open == TRUE),
      days_closed    = sum(!is.na(is_open) & is_open == FALSE),
      days_indeterminate = sum(is.na(is_open)),
      note           = first(confirmed_note),
      .groups = "drop"
    ),
  caption = "Hours overrides by facility pattern"
)
```

## Summary

```{r summary}
kable(data.frame(
  Item = c(
    "Facilities with hours overrides",
    "Total override rows (hours)",
    "Subsidy overrides"
  ),
  Count = c(
    n_distinct(hours_overrides$name_pattern),
    nrow(hours_overrides),
    nrow(subsidy_overrides)
  )
))
```

## Data Quality Audit

### Override Coverage

```{r audit-coverage}
# Check that every pattern has a full 7-day set of override rows
day_coverage <- hours_overrides |>
  group_by(name_pattern) |>
  summarise(n_days = n_distinct(day), .groups = "drop")

incomplete <- day_coverage |> filter(n_days < 7)
if (nrow(incomplete) > 0) {
  cat("WARNING: The following patterns have fewer than 7 day rows:\n")
  kable(incomplete)
} else {
  cat("All patterns have a full 7-day set of override rows.\n")
}
```

## Manual Audit Log

This section is a static, append-only record of verification calls. It does
not recompute on render.

| Date | Facility pattern | Action taken |
|---|---|---|
| 2026-03-26 | annie mae bembury | Phone call: M–F 24hr confirmed; Sat 6am–6pm confirmed; Sun by parent arrangement |
| 2026-03-26 | sylvia bennette | Phone call: M–F 6:30am–6pm confirmed |
| 2026-03-26 | pure at heart childcare | Phone call: M–F 6am–6pm confirmed; late-night/weekend services not currently operating |
| 2026-03-26 | alstons village | Phone call: M–F 6am–6pm confirmed |
| 2026-03-26 | live laugh grow | Phone call: M–F 6am–6pm confirmed |
| 2026-03-26 | kimberly glover | Voicemail left; no response received. No override applied pending confirmation. |
