---
title: "Hampton Childcare — Merge Providers"
date: "`r Sys.Date()`"
format: gfm
---

## Overview

Joins CCAV and DSS standardized data into a single master provider table using
a full outer join on `name_key`. No facility from either source is dropped.
A companion long-format hours table (one row per facility × day) is also
produced, with phone-confirmed overrides applied and per-day CCAV/DSS
discrepancies flagged.

**Source preference:** CCAV is preferred for name and program type; DSS is
preferred for address, licensure, and subsidy data.

**Input:**
- `Data/standardize/CCAV_standardize_YYYY-MM-DD.csv`
- `Data/standardize/DSS_standardize_YYYY-MM-DD.csv`
- `Data/standardize/hours_long_YYYY-MM-DD.csv`
- `Data/merge/manual_overrides_hours.csv`

**Output:**
- `Data/merge/providers_merged_YYYY-MM-DD.csv` — one row per facility
- `Data/merge/providers_hours_merged_YYYY-MM-DD.csv` — one row per facility × day

| Limitation | Detail |
|---|---|
| Join key is name_key only | Address is NOT a join key due to NAs and format drift between sources. |
| CCAV-only facilities have no subsidy data | `accepts_subsidy = NA` for any facility not in DSS. |
| Source coverage differences unexplained | The two sources have different maintainers and refresh cycles; coverage differences are treated as observed, not interpreted. |
| Duplicate name_keys produce many-to-many rows | Pipeline warns and continues; review the output if counts look wrong. |

---

```{r setup, include=FALSE}
library(here)
library(readr)
library(dplyr)
library(stringr)
library(purrr)
library(tidyr)
library(knitr)

source(here::here("config.R"))
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

today <- format(Sys.Date(), "%Y-%m-%d")
```

## Helper Functions

```{r helpers, include=FALSE}
locate_latest <- function(dir, pattern) {
  f <- list.files(dir, pattern = pattern, full.names = TRUE)
  if (!length(f)) stop("No file matching '", pattern, "' in ", dir)
  f[which.max(file.mtime(f))]
}

# Returns TRUE when two values differ (treating NA vs NA as no difference,
# NA vs non-NA as a difference)
field_differs <- function(a, b) {
  if (is.na(a) && is.na(b)) return(FALSE)
  if (is.na(a) || is.na(b)) return(TRUE)
  as.character(a) != as.character(b)
}

std_phone <- function(x) str_remove_all(x, "[^0-9]")
std_time  <- function(x) str_extract(x, "^[0-9]{2}:[0-9]{2}")
```

## Load Data

```{r load-data}
ccav_path       <- locate_latest(here("Data", "02_standardize"), "^CCAV_standardize_.*\\.csv$")
dss_path        <- locate_latest(here("Data", "02_standardize"), "^DSS_standardize_.*\\.csv$")
hours_long_path <- locate_latest(here("Data", "02_standardize"), "^hours_long_.*\\.csv$")
overrides_path  <- here("Data", "overrides", "manual_overrides_hours.csv")

cat(sprintf("CCAV:       %s\nDSS:        %s\nHours long: %s\n",
  basename(ccav_path), basename(dss_path), basename(hours_long_path)))

ccav           <- read_csv(ccav_path,       show_col_types = FALSE)
dss            <- read_csv(dss_path,        show_col_types = FALSE)
hours_long_raw <- read_csv(hours_long_path, show_col_types = FALSE)
overrides      <- read_csv(overrides_path,  show_col_types = FALSE)

out_date <- str_extract(basename(hours_long_path), "[0-9]{4}-[0-9]{2}-[0-9]{2}")
if (is.na(out_date)) out_date <- today
```

## Prepare for Merge

Rename columns with source suffixes before the join so fields from each source
remain distinguishable in the merged table.

```{r slim-sources}
ccav_slim <- ccav |>
  select(
    name_key,
    address_key_ccav    = address_key,
    name_ccav           = name,
    address_ccav        = address,
    phone_ccav          = phone,
    capacity_ccav       = capacity,
    program_type        = type,
    license_type_ccav   = licenseType,
    license_id_ccav     = licenseId,
    director,
    age_min_months_ccav = age_min_months,
    age_max_months_ccav = age_max_months,
    serves_under_5_ccav = serves_under_5,
    days_ccav           = days,
    open_time_ccav      = open_time,
    close_time_ccav     = close_time,
    ages_raw_ccav       = ages
  )

dss_slim <- dss |>
  select(
    name_key,
    address_key_dss    = address_key,
    name_dss           = name,
    address_dss        = address,
    phone_dss          = phone,
    capacity_dss       = capacity,
    facility_type      = facility_type,
    license_type_dss   = license_type,
    license_id_dss     = license_id,
    expiration_date,
    administrator,
    inspector,
    subsidy_provider,
    dss_tab            = tab,
    age_min_months_dss = age_min_months,
    age_max_months_dss = age_max_months,
    serves_under_5_dss = serves_under_5,
    days_dss           = days,
    open_time_dss      = open_time,
    close_time_dss     = close_time,
    ages_raw_dss       = ages_raw
  )
```

## Full Outer Join

```{r join}
# Warn on duplicate name_keys — these produce many-to-many rows and inflate counts
dupe_ccav <- ccav_slim |> group_by(name_key) |> filter(n() > 1) |> pull(name_key) |> unique()
dupe_dss  <- dss_slim  |> group_by(name_key) |> filter(n() > 1) |> pull(name_key) |> unique()
if (length(dupe_ccav) > 0) warning("Duplicate name_key in CCAV: ", paste(dupe_ccav, collapse = ", "))
if (length(dupe_dss)  > 0) warning("Duplicate name_key in DSS: ",  paste(dupe_dss,  collapse = ", "))

merged_raw <- full_join(ccav_slim, dss_slim, by = "name_key", relationship = "many-to-many") |>
  mutate(
    in_ccav = !is.na(name_ccav),
    in_dss  = !is.na(name_dss),
    source  = case_when(
      in_ccav & in_dss ~ "Both",
      in_ccav          ~ "CCAV only",
      TRUE             ~ "DSS only"
    )
  )
```

## Coalesce and Derive Fields

```{r coalesce}
merged <- merged_raw |>
  mutate(
    # CCAV preferred for display name and program type; DSS for everything else
    name           = coalesce(name_ccav, name_dss),
    address        = coalesce(address_dss, address_ccav),
    phone          = coalesce(phone_dss,   phone_ccav),
    capacity       = coalesce(capacity_dss, capacity_ccav),
    age_min_months = coalesce(age_min_months_dss, age_min_months_ccav),
    age_max_months = coalesce(age_max_months_dss, age_max_months_ccav),
    serves_under_5 = coalesce(serves_under_5_dss, serves_under_5_ccav),
    ages_raw       = coalesce(ages_raw_dss, ages_raw_ccav),
    address_key    = coalesce(address_key_ccav, address_key_dss)
  )
```

## Discrepancy Flags

Flag fields that differ between sources for facilities present in both. These
are not errors — they reflect different refresh cycles and maintainers.

```{r discrepancy-flags}
merged <- merged |>
  mutate(
    .disc_address  = map2_lgl(address_key_ccav,          address_key_dss,         field_differs),
    .disc_phone    = map2_lgl(std_phone(phone_ccav),     std_phone(phone_dss),    field_differs),
    .disc_capacity = map2_lgl(capacity_ccav,             capacity_dss,            field_differs),
    .disc_age_min  = map2_lgl(age_min_months_ccav,       age_min_months_dss,      field_differs),
    .disc_age_max  = map2_lgl(age_max_months_ccav,       age_max_months_dss,      field_differs),
    .disc_days     = map2_lgl(days_ccav,                 days_dss,                field_differs),
    .disc_open     = map2_lgl(std_time(open_time_ccav),  std_time(open_time_dss), field_differs),
    .disc_close    = map2_lgl(std_time(close_time_ccav), std_time(close_time_dss),field_differs)
  ) |>
  mutate(
    discrepant_fields = if_else(
      source == "Both",
      pmap_chr(
        list(.disc_address, .disc_phone, .disc_capacity,
             .disc_age_min, .disc_age_max, .disc_days, .disc_open, .disc_close),
        function(addr, ph, cap, amin, amax, dy, op, cl) {
          labels <- c("address", "phone", "capacity",
                      "age_min", "age_max", "days", "open_time", "close_time")
          diffs  <- labels[c(addr, ph, cap, amin, amax, dy, op, cl)]
          if (length(diffs) == 0) NA_character_ else paste(diffs, collapse = "; ")
        }
      ),
      NA_character_
    ),
    has_discrepancy = !is.na(discrepant_fields)
  ) |>
  select(-starts_with(".disc_"), -address_key_ccav, -address_key_dss)
```

## Derived Indicators and Subsidy Status

```{r derive-indicators}
merged <- merged |>
  mutate(
    serves_infants = !is.na(age_min_months) & age_min_months < 12,
    special_needs  = str_detect(program_type, regex("ECSE|special needs", ignore_case = TRUE)),
    # CCAV-only facilities have no DSS record → subsidy status is NA
    # No assumptions are made about subsidy status for CCAV-only facilities
    accepts_subsidy = case_when(
      str_to_lower(subsidy_provider) == "yes" ~ TRUE,
      str_to_lower(subsidy_provider) == "no"  ~ FALSE,
      TRUE                                    ~ NA
    )
  )
```

## Write Wide Facilities Output

```{r write-wide}
output_path <- here("Data", "03_merge", paste0("providers_merged_", today, ".csv"))
write_csv(merged, output_path)
prune_dated(here("Data", "03_merge"), "^providers_merged_[0-9]{4}-[0-9]{2}-[0-9]{2}\\.csv$")
cat(sprintf("Wrote %d rows → %s\n", nrow(merged), basename(output_path)))
```

## Build Long Hours Table

```{r build-hours-long}
DAY_ORDER     <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
all_name_keys <- merged$name_key

hours_ccav <- hours_long_raw |>
  filter(source == "CCAV") |>
  select(name_key, day, open_time_ccav = open_time, close_time_ccav = close_time, is_open_ccav = is_open)

hours_dss <- hours_long_raw |>
  filter(source == "DSS") |>
  select(name_key, day, open_time_dss = open_time, close_time_dss = close_time, is_open_dss = is_open)

hours_merged <- tibble(
  name_key = rep(all_name_keys, each = 7),
  day      = rep(DAY_ORDER, times = length(all_name_keys))
) |>
  left_join(hours_ccav, by = c("name_key", "day")) |>
  left_join(hours_dss,  by = c("name_key", "day")) |>
  mutate(
    open_time       = coalesce(open_time_dss,  open_time_ccav),
    close_time      = coalesce(close_time_dss, close_time_ccav),
    is_open         = coalesce(is_open_dss,    is_open_ccav),
    hours_confirmed = FALSE
  )
```

## Apply Manual Hours Overrides

Each name_pattern is matched by substring to all name_keys. The pipeline stops
if any pattern matches no facility — this prevents silent override failures.

```{r apply-overrides}
all_keys <- hours_merged |> distinct(name_key)

pattern_to_keys <- overrides |>
  distinct(name_pattern) |>
  cross_join(all_keys) |>
  filter(str_detect(name_key, fixed(name_pattern)))

unmatched_patterns <- overrides |>
  distinct(name_pattern) |>
  anti_join(pattern_to_keys, by = "name_pattern")

if (nrow(unmatched_patterns) > 0) {
  stop("Override pattern(s) matched no facilities:\n",
       paste(unmatched_patterns$name_pattern, collapse = "\n"))
}

overrides_expanded <- overrides |>
  left_join(pattern_to_keys, by = "name_pattern", relationship = "many-to-many") |>
  select(name_key, day, open_time, close_time, is_open)

hours_merged <- hours_merged |>
  left_join(
    overrides_expanded |>
      mutate(.has_ov = TRUE) |>
      rename(.open = open_time, .close = close_time, .is_open = is_open),
    by = c("name_key", "day")
  ) |>
  mutate(
    open_time       = if_else(coalesce(.has_ov, FALSE), .open,    open_time),
    close_time      = if_else(coalesce(.has_ov, FALSE), .close,   close_time),
    is_open         = if_else(coalesce(.has_ov, FALSE), .is_open, is_open),
    hours_confirmed = hours_confirmed | coalesce(.has_ov, FALSE)
  ) |>
  select(-.open, -.close, -.is_open, -.has_ov)
```

## Per-Day Discrepancy Flag

```{r hours-discrepancy}
hours_merged <- hours_merged |>
  mutate(
    time_discrepancy = case_when(
      is.na(open_time_ccav)  | is.na(open_time_dss)  ~ FALSE,
      is.na(close_time_ccav) | is.na(close_time_dss) ~ FALSE,
      open_time_ccav  != open_time_dss               ~ TRUE,
      close_time_ccav != close_time_dss              ~ TRUE,
      TRUE                                           ~ FALSE
    )
  ) |>
  select(name_key, day,
         open_time_ccav, close_time_ccav, is_open_ccav,
         open_time_dss,  close_time_dss,  is_open_dss,
         open_time, close_time, is_open,
         hours_confirmed, time_discrepancy)
```

## Write Long Hours Output

```{r write-hours}
hours_output_path <- here("Data", "03_merge", paste0("providers_hours_merged_", out_date, ".csv"))
write_csv(hours_merged, hours_output_path)
prune_dated(here("Data", "03_merge"), "^providers_hours_merged_[0-9]{4}-[0-9]{2}-[0-9]{2}\\.csv$")
cat(sprintf("Wrote %d rows → %s\n", nrow(hours_merged), basename(hours_output_path)))
cat(sprintf("  Confirmed overrides applied: %d\n", sum(hours_merged$hours_confirmed)))
cat(sprintf("  Per-day time discrepancies:  %d\n", sum(hours_merged$time_discrepancy)))
```

## Main Findings

```{r main-findings}
n_total     <- nrow(merged)
n_both      <- sum(merged$source == "Both")
n_ccav_only <- sum(merged$source == "CCAV only")
n_dss_only  <- sum(merged$source == "DSS only")

kable(data.frame(
  Metric = c(
    "Total facilities", "Matched in both sources",
    "CCAV only (no DSS record)", "DSS only (no CCAV record)",
    "Subsidy status known", "Subsidy status NA (CCAV only)",
    "Confirmed override rows applied", "Per-day time discrepancies"
  ),
  Count = c(
    n_total, n_both, n_ccav_only, n_dss_only,
    sum(!is.na(merged$accepts_subsidy)),
    sum(is.na(merged$accepts_subsidy)),
    sum(hours_merged$hours_confirmed),
    sum(hours_merged$time_discrepancy)
  )
))
```

## Summary

```{r summary-discrepancies}
disc_freq <- merged$discrepant_fields[!is.na(merged$discrepant_fields)] |>
  str_split("; ") |> unlist() |> table() |> sort(decreasing = TRUE)

cat(sprintf(
  "%d of %d matched facilities have at least one differing field.\n\n",
  sum(merged$has_discrepancy, na.rm = TRUE), n_both
))

kable(
  data.frame(Field = names(disc_freq), Discrepant_facilities = as.integer(disc_freq)),
  caption = "Fields with source discrepancies (facilities in both sources only)"
)
```

```{r summary-overrides}
kable(
  pattern_to_keys |>
    group_by(name_pattern) |>
    summarise(matched_keys = paste(sort(name_key), collapse = "; "), n = n(), .groups = "drop"),
  caption = "Override patterns and matched facilities"
)
```

## Data Quality Audit

### CCAV-only Facilities

```{r audit-ccav-only}
ccav_only_rows <- merged |>
  filter(source == "CCAV only") |>
  select(name, program_type, serves_under_5) |>
  arrange(name)

cat(sprintf("%d facilities appear in CCAV only (no DSS record — subsidy status unknown):\n\n",
  nrow(ccav_only_rows)))
kable(ccav_only_rows)
```

### DSS-only Facilities

```{r audit-dss-only}
dss_only_rows <- merged |>
  filter(source == "DSS only") |>
  select(name, dss_tab, serves_under_5) |>
  arrange(name)

cat(sprintf("%d facilities appear in DSS only (no CCAV record):\n\n", nrow(dss_only_rows)))
kable(dss_only_rows)
```

### Missing Data

```{r audit-missing}
kable(data.frame(
  Field    = c("address", "age range", "subsidy status"),
  NA_count = c(
    sum(is.na(merged$address)),
    sum(is.na(merged$age_min_months)),
    sum(is.na(merged$subsidy_provider))
  ),
  Note = c(
    sprintf("%d CCAV, %d DSS", sum(is.na(merged$address_ccav)), sum(is.na(merged$address_dss))),
    "no age data in source",
    "CCAV-only; no DSS record"
  )
))
```

### Random Spot Check

```{r audit-spot-check}
set.seed(as.integer(Sys.Date()))
kable(
  slice_sample(merged, n = min(5, nrow(merged))) |>
    select(name, source, age_min_months, serves_under_5, accepts_subsidy, has_discrepancy),
  caption = "Random date-seeded spot check"
)
```
