---
title: "Hampton Childcare — Visualizations"
date: "`r Sys.Date()`"
format: gfm
# To render HTML instead:
#   quarto render Scripts/05_visualize/visualizations.qmd --to html --output-dir Reports/05_visualize
---

## Overview

Produces five donut charts summarizing the headline metrics from the Hampton
childcare access analysis. These charts are the primary deliverable to the
City of Hampton's BHCLI Data Track mayoral team.

All charts use the under-5 analysis dataset as the universe (facilities whose
youngest served age is below 60 months, YMCA excluded). Counts and percentages
are derived directly from the analysis output CSV — no additional calculations
are performed here.

| Chart | File |
|---|---|
| Subsidy acceptance (all under-5) | `Plots/subsidy_donut.png` |
| Infant service (all under-5) | `Plots/infants_donut.png` |
| Infant service by subsidy status | `Plots/infants_subsidy_donut.png` |
| Nontraditional hours (all under-5) | `Plots/nontrad_donut.png` |
| Nontraditional hours by subsidy status | `Plots/nontrad_subsidy_donut.png` |

Nontraditional hours charts show a range (solid + hatched arc) because CCAV
and DSS hours data sometimes disagrees. The solid arc reflects the conservative
(low) estimate; the hatched extension reflects the range to the liberal (high)
estimate.

| Limitation | Detail |
|---|---|
| Unknown subsidy status | CCAV-only facilities are counted in the denominator but have no confirmed subsidy status. |
| Hours range | Nontraditional hours are a range, not a single figure, due to source discrepancies. |

---

```{r setup, include=FALSE}
library(here)
library(tidyverse)
library(ggforce)
library(ggpattern)
library(readr)

source(here::here("config.R"))
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = here::here())

# Load the most recent analysis output
analysis_file <- tail(sort(Sys.glob(here::here("Data", "04_analyze", "providers_analysis_*.csv"))), 1)
if (length(analysis_file) == 0) stop("No analysis output found in Data/analysis/")
analysis <- read_csv(analysis_file, show_col_types = FALSE)
u5 <- analysis

plots_dir <- here::here("Plots")
dir.create(plots_dir, showWarnings = FALSE, recursive = TRUE)

# Color palette — viridis-adjacent, consistent across all charts
clr_yes <- "#8856A7"  # dark purple
clr_no  <- "#9EBCDA"  # medium blue
clr_unk <- "#E0ECF4"  # light blue

# Derived counts used across charts
sub_data <- u5 |>
  mutate(status = case_when(
    accepts_subsidy == TRUE  ~ "Accepts subsidy",
    accepts_subsidy == FALSE ~ "Does not accept subsidy",
    TRUE                     ~ "Unknown"
  )) |>
  count(status)

inf_data <- u5 |>
  mutate(status = if_else(serves_infants, "Serves infants", "Does not serve\ninfants")) |>
  count(status)

inf_sub_split <- u5 |>
  mutate(status = case_when(
    serves_infants & accepts_subsidy == TRUE  ~ "Serves infants\n& accepts subsidy",
    serves_infants & accepts_subsidy == FALSE ~ "Serves infants,\ndoes not accept subsidy",
    TRUE                                      ~ "Does not serve\ninfants"
  )) |>
  count(status)

# Nontraditional hours — pre-computed low/high in analysis CSV
nt_low      <- sum(u5$nontrad_low,                               na.rm = TRUE)
nt_high     <- sum(u5$nontrad_high,                              na.rm = TRUE)
nt_sub_low  <- sum(u5$nontrad_low  & u5$accepts_subsidy == TRUE, na.rm = TRUE)
nt_sub_high <- sum(u5$nontrad_high & u5$accepts_subsidy == TRUE, na.rm = TRUE)
n_total     <- nrow(u5)
```

## Helper Functions

```{r donut-fn, include=FALSE}
# Standard donut: one slice per category with leader lines and percent labels
donut_chart <- function(df, status_col, n_col, colors, title) {
  total <- sum(df[[n_col]])
  df2 <- df |>
    mutate(
      prop        = .data[[n_col]] / total,
      end_angle   = 2 * pi * cumsum(prop),
      start_angle = lag(end_angle, default = 0),
      mid_angle   = (start_angle + end_angle) / 2,
      lx          = sin(mid_angle) * 1.45,
      ly          = cos(mid_angle) * 1.45,
      sx0         = sin(mid_angle) * 1.03,
      sy0         = cos(mid_angle) * 1.03,
      sx1         = sin(mid_angle) * 1.30,
      sy1         = cos(mid_angle) * 1.30,
      hjust       = if_else(sin(mid_angle) >= 0, 0, 1),
      label_text  = paste0(.data[[status_col]], "\n", round(100 * prop), "%")
    )

  ggplot(df2) +
    geom_arc_bar(
      aes(x0 = 0, y0 = 0, r0 = 0, r = 1,
          start = start_angle, end = end_angle,
          fill = .data[[status_col]]),
      color = "white", linewidth = 0.8
    ) +
    geom_segment(
      aes(x = sx0, y = sy0, xend = sx1, yend = sy1),
      color = "grey55", linewidth = 0.4
    ) +
    geom_text(
      aes(x = lx, y = ly, label = label_text, hjust = hjust),
      size = 5.6, lineheight = 1.0, color = "grey20"
    ) +
    scale_fill_manual(values = colors, guide = "none") +
    coord_equal(clip = "off") +
    xlim(-1.7, 1.7) + ylim(-1.5, 1.5) +
    labs(title = title) +
    theme_void() +
    theme(
      plot.title  = element_text(hjust = 0.5, size = 15, face = "bold",
                                 margin = margin(b = 12)),
      plot.margin = margin(20, 40, 20, 40)
    )
}

# Build polygon data for the hatched uncertainty band on range donuts
arc_polygon <- function(start, end, r = 1, n = 200) {
  theta <- seq(start, end, length.out = n)
  rbind(data.frame(x = 0, y = 0),
        data.frame(x = sin(theta) * r, y = cos(theta) * r))
}

# Range donut: solid arc = confirmed low estimate, hatched arc = uncertain range.
# Used for nontraditional hours where CCAV/DSS disagreement produces a range.
range_donut <- function(n_low, n_high, n_total, clr_solid, clr_rest, title,
                        label_combined, label_rest, caption = NULL) {
  n_range  <- n_high - n_low
  pct_low  <- round(100 * n_low  / n_total)
  pct_high <- round(100 * n_high / n_total)
  pct_rest <- round(100 * (n_total - n_high) / n_total)

  segs <- tibble(n = c(n_low, n_range, n_total - n_high)) |>
    mutate(
      prop        = n / n_total,
      end_angle   = 2 * pi * cumsum(prop),
      start_angle = lag(end_angle, default = 0)
    )

  df_combined <- tibble(start_angle = segs$start_angle[1], end_angle = segs$end_angle[2])
  df_rest     <- segs[3, ]
  hatch_poly  <- arc_polygon(segs$start_angle[2], segs$end_angle[2])

  combined_mid <- (segs$start_angle[1] + segs$end_angle[2]) / 2
  rest_mid     <- (segs$start_angle[3] + segs$end_angle[3]) / 2

  labels <- data.frame(
    mid = c(combined_mid, rest_mid),
    txt = c(paste0(label_combined, "\n", pct_low, "–", pct_high, "%"),
            paste0(label_rest, "\n", pct_rest, "%"))
  ) |> mutate(
    lx    = sin(mid) * 1.45,
    ly    = cos(mid) * 1.45,
    sx0   = sin(mid) * 1.03,
    sy0   = cos(mid) * 1.03,
    sx1   = sin(mid) * 1.30,
    sy1   = cos(mid) * 1.30,
    hjust = ifelse(sin(mid) >= 0, 0, 1)
  )

  ggplot() +
    geom_arc_bar(
      data = bind_rows(df_combined, df_rest) |> mutate(grp = c("a", "b")),
      aes(x0 = 0, y0 = 0, r0 = 0, r = 1,
          start = start_angle, end = end_angle, fill = grp),
      color = "white", linewidth = 0.8
    ) +
    geom_polygon_pattern(
      data            = hatch_poly,
      aes(x = x, y = y),
      fill            = NA,
      pattern         = "stripe",
      pattern_colour  = "white",
      pattern_fill    = "white",
      pattern_angle   = 35,
      pattern_density = 0.1,
      pattern_spacing = 0.007,
      color           = NA
    ) +
    geom_segment(
      data = labels,
      aes(x = sx0, y = sy0, xend = sx1, yend = sy1),
      color = "grey55", linewidth = 0.4
    ) +
    geom_text(
      data = labels,
      aes(x = lx, y = ly, label = txt, hjust = hjust),
      size = 5.6, lineheight = 1.0, color = "grey20"
    ) +
    scale_fill_manual(values = c("a" = clr_solid, "b" = clr_rest), guide = "none") +
    coord_equal(clip = "off") +
    xlim(-1.7, 1.7) + ylim(-1.5, 1.5) +
    labs(title = title, caption = caption) +
    theme_void() +
    theme(
      plot.title            = element_text(hjust = 0.5, size = 15, face = "bold",
                                           margin = margin(b = 12)),
      plot.caption          = element_text(hjust = 1, size = 9, color = "grey50",
                                           margin = margin(t = 10)),
      plot.caption.position = "plot",
      plot.margin           = margin(20, 40, 30, 40)
    )
}
```

## Charts

```{r subsidy-donut, fig.width=8, fig.height=6}
p1 <- donut_chart(
  sub_data, "status", "n",
  colors = c("Accepts subsidy"         = clr_yes,
             "Does not accept subsidy" = clr_no,
             "Unknown"                 = clr_unk),
  title  = paste0("Under-5 Facilities that Accept Subsidy\n(n = ", n_total, ")")
)
p1
ggsave(file.path(plots_dir, "subsidy_donut.png"),
       p1, width = 8, height = 6, dpi = 150, bg = "white")
```

```{r infants-donut, fig.width=8, fig.height=6}
p2 <- donut_chart(
  inf_data, "status", "n",
  colors = c("Serves infants"          = clr_yes,
             "Does not serve\ninfants" = clr_no),
  title  = paste0("Under-5 Facilities that Serve Infants\n(n = ", n_total, ")")
)
p2
ggsave(file.path(plots_dir, "infants_donut.png"),
       p2, width = 8, height = 6, dpi = 150, bg = "white")
```

```{r infants-subsidy-donut, fig.width=8, fig.height=6}
p3 <- donut_chart(
  inf_sub_split, "status", "n",
  colors = c("Serves infants\n& accepts subsidy"        = clr_yes,
             "Serves infants,\ndoes not accept subsidy" = clr_unk,
             "Does not serve\ninfants"                  = clr_no),
  title  = paste0("Under-5 Facilities that Serve Infants by Subsidy Status\n(n = ", n_total, ")")
)
p3
ggsave(file.path(plots_dir, "infants_subsidy_donut.png"),
       p3, width = 8, height = 6, dpi = 150, bg = "white")
```

```{r nontrad-donut, fig.width=8, fig.height=6}
p4 <- range_donut(
  n_low          = nt_low,
  n_high         = nt_high,
  n_total        = n_total,
  clr_solid      = clr_yes,
  clr_rest       = clr_no,
  title          = paste0("Under-5 Facilities Open Non-traditional Hours\n(8 pm–5 am,  n = ", n_total, ")"),
  label_combined = "Non-traditional hours",
  label_rest     = "Traditional hours",
  caption        = "Range reflects source discrepancies"
)
p4
ggsave(file.path(plots_dir, "nontrad_donut.png"),
       p4, width = 8, height = 6, dpi = 150, bg = "white")
```

```{r nontrad-subsidy-donut, fig.width=8, fig.height=6}
p5 <- range_donut(
  n_low          = nt_sub_low,
  n_high         = nt_sub_high,
  n_total        = n_total,
  clr_solid      = clr_yes,
  clr_rest       = clr_no,
  title          = paste0("Under-5 Facilities Open Non-traditional Hours by Subsidy Status\n(8 pm–5 am,  n = ", n_total, ")"),
  label_combined = "Non-traditional hours\n& accepts subsidy",
  label_rest     = "Traditional hours",
  caption        = "Range reflects source discrepancies"
)
p5
ggsave(file.path(plots_dir, "nontrad_subsidy_donut.png"),
       p5, width = 8, height = 6, dpi = 150, bg = "white")
```

## Data Quality Audit

### Input Verification

```{r audit-input}
cat(sprintf("Analysis file: %s\n", basename(analysis_file)))
cat(sprintf("Rows loaded:   %d\n", nrow(u5)))
cat(sprintf("Expected columns present: %s\n",
  all(c("nontrad_low", "nontrad_high", "serves_infants", "accepts_subsidy") %in% names(u5))))
```

### Count Cross-Check

```{r audit-counts}
knitr::kable(data.frame(
  Chart = c("Subsidy", "Infants", "Infants × subsidy", "Nontrad (low)", "Nontrad (high)"),
  n_plotted = c(
    sum(sub_data$n),
    sum(inf_data$n),
    sum(inf_sub_split$n),
    nt_low + (n_total - nt_high) + (nt_high - nt_low),
    n_total
  ),
  n_expected = n_total
), caption = "All chart denominators should equal n_total")
```

### Plot Files Written

```{r audit-plots}
plot_files <- c(
  "subsidy_donut.png", "infants_donut.png", "infants_subsidy_donut.png",
  "nontrad_donut.png", "nontrad_subsidy_donut.png"
)
knitr::kable(data.frame(
  File   = plot_files,
  Exists = file.exists(file.path(plots_dir, plot_files)),
  Size_kb = round(file.size(file.path(plots_dir, plot_files)) / 1024, 1)
), caption = "Generated plot files")
```
