---
title: "Hampton Childcare — CCAV Standardize"
date: "`r Sys.Date()`"
format: gfm
---

## Overview

Cleans and normalizes the CCAV ingest data to produce merge-ready fields. The
two key transformations are age range parsing (free-text string → numeric
months) and hours parsing (day-and-time string → structured `days`,
`open_time`, `close_time` fields). A name key and address key are also
generated for the downstream join with DSS.

**Input:** `Data/ingest/CCAV_ingest_YYYY-MM-DD.csv`
**Output:** `Data/standardize/CCAV_standardize_YYYY-MM-DD.csv`

| Step | Input format | Output |
|---|---|---|
| Address cleaning | `"N Director Name 123 Real St"` | Street-only address |
| Age range parsing | `"1 Year 4 Months - 12 Years 11 Months"` | `age_min_months`, `age_max_months` |
| Hours parsing | `"Monday - Friday 6:00am - 8:00pm"` | `days`, `open_time`, `close_time` (24h HH:MM) |
| Name key | Honorifics stripped, lowercased, punctuation removed | `name_key` |
| Address key | Uppercased, `"City VA XXXXX"` → `"City, VA XXXXX"` | `address_key` |

| Limitation | Detail |
|---|---|
| Age data absent for some facilities | Parsed values will be NA; those facilities are excluded in the analysis stage. |
| Hours format variation | CCAV hours strings vary; unrecognized formats produce NA open/close times. |
| Address prefix contamination | Some addresses begin with a director name; the director field is used to strip it. |

---

```{r setup, include=FALSE}
library(here)
library(readr)
library(dplyr)
library(stringr)
library(purrr)
library(knitr)

source(here::here("config.R"))
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

today <- format(Sys.Date(), "%Y-%m-%d")
```

## Helper Functions

```{r helpers, include=FALSE}
# Parse a single age bound string (e.g. "1 Year 4 Months") to total months.
# Extracts the first year and month value found; coerces missing to 0.
parse_age_to_months <- function(x) {
  x <- str_to_lower(str_squish(x))
  years  <- coalesce(as.numeric(str_extract(x, "(\\d+)(?=\\s*year)")),  0)
  months <- coalesce(as.numeric(str_extract(x, "(\\d+)(?=\\s*month)")), 0)
  years * 12 + months
}

parse_age_range <- function(age_str) {
  parts <- str_split(age_str, "\\s*-\\s*", n = 2)[[1]]
  if (length(parts) < 2) {
    return(list(age_min_months = NA_real_, age_max_months = NA_real_))
  }
  list(
    age_min_months = parse_age_to_months(parts[1]),
    age_max_months = parse_age_to_months(parts[2])
  )
}

# Parse CCAV hours string to days range, open time, close time (24h HH:MM).
# CCAV format: "Monday - Friday 6:00am - 8:00pm"
parse_ccav_hours <- function(h) {
  h <- str_squish(h)

  # Day portion is everything before the first digit; time portion is from first digit
  day_part  <- str_extract(h, "^[^\\d]+(?=\\d)") |> str_squish()
  time_part <- str_extract(h, "\\d.*$")           |> str_squish()

  day_range <- str_replace_all(day_part, "\\s*-\\s*", "-") |>
    str_remove("-$") |>
    str_squish()

  times <- str_extract_all(time_part, "\\d{1,2}:\\d{2}", simplify = FALSE)[[1]]
  ampm  <- str_extract_all(time_part, "(?<=\\d{2})(am|pm)", simplify = FALSE)[[1]]

  to_24h <- function(t, ap) {
    if (is.na(t) || t == "") return(NA_character_)
    parts <- str_split(t, ":")[[1]]
    h <- as.integer(parts[1])
    m <- as.integer(parts[2])
    if (!is.na(ap)) {
      ap <- str_to_lower(ap)
      if (ap == "pm" && h != 12) h <- h + 12
      if (ap == "am" && h == 12) h <- 0L
    }
    sprintf("%02d:%02d", h, m)
  }

  open_time  <- if (length(times) >= 1) to_24h(times[1], ampm[1]) else NA_character_
  close_time <- if (length(times) >= 2) to_24h(times[2], ampm[2]) else NA_character_

  list(days = day_range, open_time = open_time, close_time = close_time)
}

# Lowercase, strip honorifics, remove punctuation — used as join key with DSS
standardise_name <- function(x) {
  x |>
    str_remove_all("^(Ms\\.|Mrs\\.|Mr\\.|Dr\\.)\\s*") |>
    str_to_lower() |>
    str_replace_all("[^a-z0-9 ]", "") |>
    str_squish()
}

# Uppercase, normalize spacing, insert missing comma before "VA XXXXX"
standardise_address <- function(x) {
  x |>
    str_replace("(?<=[a-zA-Z])\\s+(VA\\s+\\d{5})", ", \\1") |>
    str_squish() |>
    str_to_upper()
}

# Strip the director name prefix from the address field.
# CCAV scraper sometimes captures "N Director Name 123 Real St"; use the
# director column to find and remove that prefix. Takes the LAST match to
# handle doubled entries from the scraper.
strip_director_prefix <- function(address, director) {
  if (is.na(address) || is.na(director) || director == "") return(address)
  locs <- str_locate_all(address, fixed(director, ignore_case = TRUE))[[1]]
  if (nrow(locs) == 0) return(address)
  str_squish(str_sub(address, max(locs[, "end"]) + 1))
}
```

## Load Data

```{r load-data}
ingest_files <- list.files(
  here("Data", "01_ingest"),
  pattern    = "^CCAV_ingest_[0-9]{4}-[0-9]{2}-[0-9]{2}\\.csv$",
  full.names = TRUE
)
if (length(ingest_files) == 0) stop("No CCAV ingest file found in Data/ingest/")
ingest_path <- ingest_files[which.max(file.mtime(ingest_files))]
cat(sprintf("Reading: %s (%d rows)\n", basename(ingest_path), nrow(read_csv(ingest_path, show_col_types = FALSE, n_max = 0))))
raw <- read_csv(ingest_path, show_col_types = FALSE)
```

## Apply Transformations

```{r transform}
processed <- raw |>
  mutate(
    address        = map2_chr(address, director, strip_director_prefix),
    age_parsed     = map(ages, parse_age_range),
    age_min_months = map_dbl(age_parsed, "age_min_months"),
    age_max_months = map_dbl(age_parsed, "age_max_months"),
    serves_under_5 = age_min_months < 60,
    hours_parsed   = map(hours, parse_ccav_hours),
    days           = map_chr(hours_parsed, "days"),
    open_time      = map_chr(hours_parsed, "open_time"),
    close_time     = map_chr(hours_parsed, "close_time"),
    name_key       = standardise_name(name),
    address_key    = standardise_address(address),
    source         = "CCAV"
  ) |>
  select(-age_parsed, -hours_parsed)
```

## Write Output

```{r write-output}
output_path <- here("Data", "02_standardize", paste0("CCAV_standardize_", today, ".csv"))
dir.create(here("Data", "02_standardize"), showWarnings = FALSE, recursive = TRUE)
write_csv(processed, output_path)
prune_dated(here("Data", "02_standardize"), "^CCAV_standardize_[0-9]{4}-[0-9]{2}-[0-9]{2}\\.csv$")
cat(sprintf("Wrote %d rows → %s\n", nrow(processed), basename(output_path)))
```

## Main Findings

```{r main-findings}
kable(data.frame(
  Metric = c(
    "Input rows", "Output rows",
    "Serves under-5 (age_min < 60 months)",
    "Serves infants (age_min < 12 months)",
    "Missing age data",
    "Missing open_time"
  ),
  Count = c(
    nrow(raw), nrow(processed),
    sum(processed$serves_under_5, na.rm = TRUE),
    sum(processed$age_min_months < 12, na.rm = TRUE),
    sum(is.na(processed$age_min_months)),
    sum(is.na(processed$open_time))
  )
))
```

## Summary

```{r summary-days}
kable(
  processed |> count(days, sort = TRUE),
  caption = "Days distribution (top values)"
)
```

## Data Quality Audit

### Age Range Distribution

```{r audit-age}
cat("age_min_months:\n")
print(summary(processed$age_min_months))
cat("\nage_max_months:\n")
print(summary(processed$age_max_months))
```

### Hours Parsing Failures

```{r audit-hours}
# Facilities with hours text but no parsed open_time indicate a format not
# handled by parse_ccav_hours — review the raw hours strings below
unparsed <- processed |>
  filter(!is.na(hours), hours != "", is.na(open_time))

if (nrow(unparsed) > 0) {
  cat(sprintf("WARNING: %d facilities have hours text but no parsed open_time:\n", nrow(unparsed)))
  kable(unparsed |> select(name, hours, days, open_time, close_time))
} else {
  cat("All facilities with hours text were successfully parsed.\n")
}
```

### Missing Fields

```{r audit-missing}
kable(data.frame(
  Field     = c("address_key", "age_min_months", "open_time", "close_time", "name_key"),
  NA_count  = c(
    sum(is.na(processed$address_key)),
    sum(is.na(processed$age_min_months)),
    sum(is.na(processed$open_time)),
    sum(is.na(processed$close_time)),
    sum(is.na(processed$name_key))
  )
))
```

### Random Spot Check

```{r audit-spot-check}
set.seed(as.integer(Sys.Date()))
kable(
  slice_sample(processed, n = min(5, nrow(processed))) |>
    select(name, age_min_months, age_max_months, days, open_time, close_time, name_key),
  caption = "Random date-seeded spot check"
)
```
