---
title: "Hampton Childcare — DSS Standardize"
date: "`r Sys.Date()`"
format: gfm
---

## Overview

Cleans and normalizes the DSS ingest data to produce merge-ready fields,
applying the same age range and hours parsing as the CCAV standardize module.
DSS hours strings use a different format (`"6:00 am - 5:00 pm, Monday -
Friday"`) and require separate parsing logic.

**Input:** `Data/ingest/DSS_ingest_YYYY-MM-DD.csv`
**Output:** `Data/standardize/DSS_standardize_YYYY-MM-DD.csv`

| Step | Input format | Output |
|---|---|---|
| Age range parsing | `"1 month - 12 years 11 months"` | `age_min_months`, `age_max_months` |
| Hours parsing | `"6:00 am - 5:00 pm, Monday - Friday"` | `days`, `open_time`, `close_time` (24h HH:MM) |
| Name key | Honorifics stripped, lowercased, punctuation removed | `name_key` |
| Address key | Uppercased, whitespace normalized | `address_key` |

| Limitation | Detail |
|---|---|
| Age data absent for some facilities | Parsed values will be NA; those facilities are excluded in the analysis stage. |
| Hours format variation | DSS hours strings vary including `a.m.`/`p.m.` variants; unrecognized formats produce NA. |
| Parse errors from ingest | Facilities flagged `parse_error = TRUE` in ingest will have mostly NA fields. |

---

```{r setup, include=FALSE}
library(here)
library(readr)
library(dplyr)
library(stringr)
library(purrr)
library(knitr)

source(here::here("config.R"))
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

today <- format(Sys.Date(), "%Y-%m-%d")
```

## Helper Functions

```{r helpers, include=FALSE}
parse_age_to_months <- function(x) {
  x <- str_to_lower(str_squish(x))
  years  <- coalesce(as.numeric(str_extract(x, "(\\d+)(?=\\s*year)")),  0)
  months <- coalesce(as.numeric(str_extract(x, "(\\d+)(?=\\s*month)")), 0)
  years * 12 + months
}

parse_age_range <- function(age_str) {
  if (is.na(age_str) || age_str == "") {
    return(list(age_min_months = NA_real_, age_max_months = NA_real_))
  }
  parts <- str_split(age_str, "\\s*-\\s*", n = 2)[[1]]
  if (length(parts) < 2) {
    return(list(age_min_months = NA_real_, age_max_months = NA_real_))
  }
  list(
    age_min_months = parse_age_to_months(parts[1]),
    age_max_months = parse_age_to_months(parts[2])
  )
}

# Normalize a.m./p.m. variants to am/pm before time extraction
normalise_ampm <- function(x) {
  x |>
    str_to_lower() |>
    str_replace_all("a\\.m\\.", "am") |>
    str_replace_all("p\\.m\\.", "pm") |>
    str_squish()
}

to_24h <- function(t, ap) {
  if (is.na(t) || t == "") return(NA_character_)
  parts <- str_split(t, ":")[[1]]
  h <- as.integer(parts[1])
  m <- as.integer(parts[2])
  if (!is.na(ap) && ap != "") {
    if (ap == "pm" && h != 12) h <- h + 12
    if (ap == "am" && h == 12) h <- 0L
  }
  sprintf("%02d:%02d", h, m)
}

# Parse DSS hours string format: "6:00 am - 5:00 pm, Monday - Friday"
# Falls back to day-name detection when no comma separator is present
parse_dss_hours <- function(h) {
  if (is.na(h) || h == "") {
    return(list(days = NA_character_, open_time = NA_character_, close_time = NA_character_))
  }

  h_norm    <- normalise_ampm(h)
  comma_pos <- str_locate(h_norm, ",")[1, "start"]

  if (!is.na(comma_pos)) {
    time_part <- str_sub(h_norm, 1, comma_pos - 1) |> str_squish()
    day_part  <- str_sub(h_norm, comma_pos + 1)    |> str_squish()
  } else {
    day_match <- str_locate(
      h_norm, "(monday|tuesday|wednesday|thursday|friday|saturday|sunday)"
    )[1, "start"]
    if (!is.na(day_match)) {
      time_part <- str_sub(h_norm, 1, day_match - 1) |> str_squish()
      day_part  <- str_sub(h_norm, day_match)         |> str_squish()
    } else {
      time_part <- h_norm
      day_part  <- NA_character_
    }
  }

  day_range <- day_part |>
    str_replace_all("\\s*-\\s*", "-") |>
    str_to_title() |>
    str_squish()

  times <- str_extract_all(time_part, "\\d{1,2}:\\d{2}")[[1]]
  ampm  <- str_extract_all(time_part, "(am|pm)")[[1]]

  open_time  <- if (length(times) >= 1) to_24h(times[1], ampm[1]) else NA_character_
  close_time <- if (length(times) >= 2) to_24h(times[2], ampm[2]) else NA_character_

  list(days = day_range, open_time = open_time, close_time = close_time)
}

standardise_name <- function(x) {
  x |>
    str_remove_all("^(Ms\\.|Mrs\\.|Mr\\.|Dr\\.)\\s*") |>
    str_to_lower() |>
    str_replace_all("[^a-z0-9 ]", "") |>
    str_squish()
}

# DSS addresses are already well-formed; uppercase and normalize spacing
standardise_address <- function(x) {
  x |>
    str_squish() |>
    str_to_upper() |>
    str_replace_all("\\s*,\\s*", ", ") |>
    str_squish()
}
```

## Load Data

```{r load-data}
ingest_files <- list.files(
  here("Data", "01_ingest"),
  pattern    = "^DSS_ingest_[0-9]{4}-[0-9]{2}-[0-9]{2}\\.csv$",
  full.names = TRUE
)
if (length(ingest_files) == 0) stop("No DSS ingest file found in Data/ingest/")
ingest_path <- ingest_files[which.max(file.mtime(ingest_files))]
cat(sprintf("Reading: %s\n", basename(ingest_path)))
raw <- read_csv(ingest_path, show_col_types = FALSE)
```

## Apply Transformations

```{r transform}
processed <- raw |>
  mutate(
    age_parsed     = map(ages_raw, parse_age_range),
    age_min_months = map_dbl(age_parsed, "age_min_months"),
    age_max_months = map_dbl(age_parsed, "age_max_months"),
    serves_under_5 = age_min_months < 60,
    hours_parsed   = map(hours, parse_dss_hours),
    days           = map_chr(hours_parsed, "days"),
    open_time      = map_chr(hours_parsed, "open_time"),
    close_time     = map_chr(hours_parsed, "close_time"),
    name_key       = standardise_name(name),
    address_key    = standardise_address(address),
    source         = "DSS"
  ) |>
  select(-age_parsed, -hours_parsed)
```

## Write Output

```{r write-output}
output_path <- here("Data", "02_standardize", paste0("DSS_standardize_", today, ".csv"))
dir.create(here("Data", "02_standardize"), showWarnings = FALSE, recursive = TRUE)
write_csv(processed, output_path)
prune_dated(here("Data", "02_standardize"), "^DSS_standardize_[0-9]{4}-[0-9]{2}-[0-9]{2}\\.csv$")
cat(sprintf("Wrote %d rows → %s\n", nrow(processed), basename(output_path)))
```

## Main Findings

```{r main-findings}
kable(data.frame(
  Metric = c(
    "Input rows", "Output rows",
    "Licensed", "Unlicensed",
    "Serves under-5 (age_min < 60 months)",
    "Subsidy providers",
    "Parse errors from ingest",
    "Missing age data",
    "Missing open_time"
  ),
  Count = c(
    nrow(raw), nrow(processed),
    sum(raw$tab == "Licensed",   na.rm = TRUE),
    sum(raw$tab == "Unlicensed", na.rm = TRUE),
    sum(processed$serves_under_5,                    na.rm = TRUE),
    sum(str_to_lower(processed$subsidy_provider) == "yes", na.rm = TRUE),
    sum(processed$parse_error, na.rm = TRUE),
    sum(is.na(processed$age_min_months)),
    sum(is.na(processed$open_time))
  )
))
```

## Summary

```{r summary-subsidy}
kable(
  processed |> count(subsidy_provider, sort = TRUE),
  caption = "Subsidy provider values (raw DSS field)"
)
```

```{r summary-days}
kable(
  processed |> count(days, sort = TRUE),
  caption = "Days distribution"
)
```

## Data Quality Audit

### Age Range Distribution

```{r audit-age}
cat("age_min_months:\n")
print(summary(processed$age_min_months))
cat("\nage_max_months:\n")
print(summary(processed$age_max_months))
```

### Hours Parsing Failures

```{r audit-hours}
unparsed <- processed |>
  filter(!is.na(hours), hours != "", is.na(open_time))

if (nrow(unparsed) > 0) {
  cat(sprintf("WARNING: %d facilities have hours text but no parsed open_time:\n", nrow(unparsed)))
  kable(unparsed |> select(name, hours, days, open_time, close_time))
} else {
  cat("All facilities with hours text were successfully parsed.\n")
}
```

### Missing Fields

```{r audit-missing}
kable(data.frame(
  Field    = c("address_key", "age_min_months", "open_time", "close_time", "name_key"),
  NA_count = c(
    sum(is.na(processed$address_key)),
    sum(is.na(processed$age_min_months)),
    sum(is.na(processed$open_time)),
    sum(is.na(processed$close_time)),
    sum(is.na(processed$name_key))
  )
))
```

### Parse Error Rows

```{r audit-parse-errors}
errors <- processed |> filter(parse_error == TRUE)
if (nrow(errors) > 0) {
  cat(sprintf("%d facilities had ingest parse errors (mostly NA fields):\n", nrow(errors)))
  kable(errors |> select(name, tab, address, hours, ages_raw))
} else {
  cat("No parse errors from ingest.\n")
}
```

### Random Spot Check

```{r audit-spot-check}
set.seed(as.integer(Sys.Date()))
kable(
  slice_sample(processed, n = min(5, nrow(processed))) |>
    select(name, age_min_months, age_max_months, days, open_time, close_time, subsidy_provider),
  caption = "Random date-seeded spot check"
)
```
