---
title: "Hampton Childcare — Hours Standardize"
date: "`r Sys.Date()`"
format: gfm
---

## Overview

Expands per-facility hours data from both standardized sources into long
format: one row per facility × day × source. This structure allows downstream
modules to compare CCAV and DSS hours side by side and to apply per-day manual
overrides precisely.

A days string like `"Monday-Friday"` is expanded to 7 rows, with `is_open =
TRUE` for Monday through Friday and `is_open = FALSE` for Saturday and Sunday.
Facilities with `NA` days get `is_open = NA` for all seven rows.

**Input:**
- `Data/standardize/CCAV_standardize_YYYY-MM-DD.csv`
- `Data/standardize/DSS_standardize_YYYY-MM-DD.csv`

**Output:** `Data/standardize/hours_long_YYYY-MM-DD.csv`

| Limitation | Detail |
|---|---|
| Facilities with NA days | Expanded with `is_open = NA` for all days — not the same as known-closed. |
| Single-day strings | Handled (e.g., `"Saturday"`). Multi-day non-range strings are not (e.g., `"Monday, Wednesday"`). |
| Open/close times copied only to open days | Days with `is_open = FALSE` get NA times regardless of source. |

---

```{r setup, include=FALSE}
library(here)
library(readr)
library(dplyr)
library(purrr)
library(stringr)
library(knitr)

source(here::here("config.R"))
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

DAY_ORDER <- c("Monday", "Tuesday", "Wednesday", "Thursday",
               "Friday", "Saturday", "Sunday")
```

## Helper Functions

```{r helpers, include=FALSE}
locate_latest <- function(dir, pattern) {
  f <- list.files(dir, pattern = pattern, full.names = TRUE)
  if (!length(f)) stop("No file matching '", pattern, "' in ", dir)
  f[which.max(file.mtime(f))]
}

# Parse a days string to a 7-element logical vector aligned to DAY_ORDER.
# Handles "Monday-Friday" range and single-day "Saturday" forms.
# Unrecognized strings produce rep(NA, 7) with a warning.
parse_days <- function(days_raw) {
  if (is.na(days_raw)) return(rep(NA, 7))

  cleaned    <- str_trim(str_remove(days_raw, "[^a-zA-Z]+$"))
  range_ends <- str_match(cleaned, "^(\\w+)-(\\w+)$")

  if (!is.na(range_ends[1, 1])) {
    start_idx <- match(range_ends[1, 2], DAY_ORDER)
    end_idx   <- match(range_ends[1, 3], DAY_ORDER)
    if (is.na(start_idx) || is.na(end_idx)) {
      warning("Unrecognised day in range: '", days_raw, "' — treating as NA")
      return(rep(NA, 7))
    }
    return(seq_along(DAY_ORDER) >= start_idx & seq_along(DAY_ORDER) <= end_idx)
  }

  single_idx <- match(cleaned, DAY_ORDER)
  if (!is.na(single_idx)) {
    result              <- rep(FALSE, 7)
    result[single_idx]  <- TRUE
    return(result)
  }

  warning("Unrecognised days value: '", days_raw, "' — treating as NA")
  rep(NA, 7)
}

# Expand one facility row to 7 rows (one per day)
expand_to_long <- function(name_key, name, source, days_raw, open_time, close_time) {
  is_open_vec <- parse_days(days_raw)
  tibble(
    name_key   = name_key,
    name       = name,
    source     = source,
    day        = DAY_ORDER,
    is_open    = is_open_vec,
    open_time  = if_else(is_open_vec %in% TRUE, as.character(open_time),  NA_character_),
    close_time = if_else(is_open_vec %in% TRUE, as.character(close_time), NA_character_),
    days_raw   = days_raw
  )
}
```

## Load Data

```{r load-data}
ccav_path <- locate_latest(here("Data", "02_standardize"), "^CCAV_standardize_.*\\.csv$")
dss_path  <- locate_latest(here("Data", "02_standardize"), "^DSS_standardize_.*\\.csv$")

ccav_date <- str_extract(basename(ccav_path), "[0-9]{4}-[0-9]{2}-[0-9]{2}")
if (is.na(ccav_date)) ccav_date <- format(Sys.Date(), "%Y-%m-%d")

cat(sprintf("CCAV: %s\nDSS:  %s\n", basename(ccav_path), basename(dss_path)))

ccav <- read_csv(ccav_path, show_col_types = FALSE)
dss  <- read_csv(dss_path,  show_col_types = FALSE)
```

## Expand to Long Format

```{r expand-ccav}
ccav_long <- map_dfr(seq_len(nrow(ccav)), function(i) {
  expand_to_long(
    name_key   = ccav$name_key[i],
    name       = ccav$name[i],
    source     = "CCAV",
    days_raw   = ccav$days[i],
    open_time  = ccav$open_time[i],
    close_time = ccav$close_time[i]
  )
})
```

```{r expand-dss}
dss_long <- map_dfr(seq_len(nrow(dss)), function(i) {
  expand_to_long(
    name_key   = dss$name_key[i],
    name       = dss$name[i],
    source     = "DSS",
    days_raw   = dss$days[i],
    open_time  = dss$open_time[i],
    close_time = dss$close_time[i]
  )
})
```

```{r bind}
hours_long <- bind_rows(ccav_long, dss_long) |>
  mutate(day = factor(day, levels = DAY_ORDER, ordered = TRUE)) |>
  arrange(source, name_key, day) |>
  mutate(day = as.character(day))  # back to character for CSV portability
```

## Write Output

```{r write-output}
output_path <- here("Data", "02_standardize", paste0("hours_long_", ccav_date, ".csv"))
write_csv(hours_long, output_path)
prune_dated(here("Data", "02_standardize"), "^hours_long_[0-9]{4}-[0-9]{2}-[0-9]{2}\\.csv$")
cat(sprintf("Wrote %d rows → %s\n", nrow(hours_long), basename(output_path)))
```

## Main Findings

```{r main-findings}
n_ccav_fac <- n_distinct(ccav_long$name_key)
n_dss_fac  <- n_distinct(dss_long$name_key)

kable(data.frame(
  Source = c("CCAV", "DSS", "Combined"),
  Facilities = c(n_ccav_fac, n_dss_fac, n_distinct(hours_long$name_key)),
  Rows = c(nrow(ccav_long), nrow(dss_long), nrow(hours_long))
))
```

## Summary

```{r summary-open}
kable(
  hours_long |>
    group_by(source, is_open) |>
    summarise(n = n(), .groups = "drop"),
  caption = "is_open distribution by source"
)
```

## Data Quality Audit

### Unrecognized Day Strings

```{r audit-days}
# Facilities where days_raw was present but is_open could not be determined
# for any day indicate a format not handled by parse_days
unrecognized <- hours_long |>
  filter(!is.na(days_raw), days_raw != "") |>
  group_by(name_key, source, days_raw) |>
  summarise(all_na = all(is.na(is_open)), .groups = "drop") |>
  filter(all_na)

if (nrow(unrecognized) > 0) {
  cat(sprintf("WARNING: %d facility × source combinations with unrecognized day strings:\n",
              nrow(unrecognized)))
  kable(unrecognized)
} else {
  cat("All day strings were recognized.\n")
}
```

### Missing Hours by Day

```{r audit-missing-hours}
kable(
  hours_long |>
    filter(coalesce(is_open, FALSE)) |>
    group_by(day) |>
    summarise(
      n_open        = n(),
      missing_open  = sum(is.na(open_time)),
      missing_close = sum(is.na(close_time)),
      .groups = "drop"
    ),
  caption = "Open days with missing open/close times"
)
```

### Random Spot Check

```{r audit-spot-check}
set.seed(as.integer(Sys.Date()))
kable(
  hours_long |>
    filter(coalesce(is_open, FALSE)) |>
    slice_sample(n = min(10, sum(coalesce(hours_long$is_open, FALSE)))) |>
    select(name_key, source, day, is_open, open_time, close_time),
  caption = "Random date-seeded spot check (open days only)"
)
```
