---
title: "Hampton Childcare — CCAV Ingest Report"
date: "`r Sys.Date()`"
format: gfm
---

## Overview

Reads the most recent CCAV ingest CSV from `Data/ingest/` and produces a
quality report. The CSV is downloaded manually using the browser console script
`Scripts/01_ingest/ccav_ingest.js` — a reCAPTCHA on the CCAV site prevents
full automation.

**Source:** ChildCare Aware of Virginia (CCAV) search tool at
`https://stage.worklifesystems.com/parent/25`, filtered to Hampton providers.

**Input:** `Data/ingest/CCAV_ingest_YYYY-MM-DD.csv`

| Limitation | Detail |
|---|---|
| Manual collection required | reCAPTCHA blocks automation; script must be run in Chrome DevTools. |
| No subsidy data | CCAV does not report subsidy acceptance. That field comes entirely from DSS. |
| Address prefix contamination | Scraper sometimes captures a director name before the street address. Cleaned in `02_standardize/ccav_standardize.qmd`. |
| Type field concatenated | Raw `type` field joins program type with offer tags. Cleaned in standardize. |

---

```{r setup, include=FALSE}
library(here)
library(readr)
library(dplyr)
library(stringr)
library(knitr)

source(here::here("config.R"))
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

## Load Data

```{r load-data}
ingest_files <- list.files(
  here("Data", "01_ingest"),
  pattern    = "^CCAV_ingest_[0-9]{4}-[0-9]{2}-[0-9]{2}\\.csv$",
  full.names = TRUE
)
if (length(ingest_files) == 0) {
  stop(
    "No CCAV ingest file found in Data/ingest/. ",
    "Run ccav_ingest.js in Chrome DevTools and move the downloaded CSV to Data/ingest/ before rendering this report."
  )
}

# Use the most recently modified file if multiple exist
ingest_path <- ingest_files[which.max(file.mtime(ingest_files))]
ingest_date <- str_extract(basename(ingest_path), "[0-9]{4}-[0-9]{2}-[0-9]{2}")
cat(sprintf("Reading: %s\n", basename(ingest_path)))

raw <- read_csv(ingest_path, show_col_types = FALSE)

# Prune old CCAV ingest CSVs — keep two most recent
prune_dated(here("Data", "01_ingest"), "^CCAV_ingest_[0-9]{4}-[0-9]{2}-[0-9]{2}\\.csv$")
```

## Main Findings

```{r main-findings}
n_total      <- nrow(raw)
n_licensed   <- sum(
  str_detect(raw$licenseType, regex("licensed", ignore_case = TRUE)) &
  !str_detect(raw$licenseType, regex("unlicensed", ignore_case = TRUE)),
  na.rm = TRUE
)
n_unlicensed <- sum(
  str_detect(raw$licenseType, regex("unlicensed|exempt|registered", ignore_case = TRUE)),
  na.rm = TRUE
)
n_missing_address <- sum(is.na(raw$address) | raw$address == "")
n_missing_hours   <- sum(is.na(raw$hours)   | raw$hours   == "")
n_missing_ages    <- sum(is.na(raw$ages)    | raw$ages    == "")

# Addresses not starting with a digit indicate a director-name prefix
n_address_prefix <- sum(
  !is.na(raw$address) & raw$address != "" & !str_detect(raw$address, "^\\d"),
  na.rm = TRUE
)

kable(data.frame(
  Metric = c(
    "Total facilities", "Licensed", "Unlicensed / exempt",
    "Missing address", "Missing hours", "Missing ages"
  ),
  Count = c(n_total, n_licensed, n_unlicensed,
            n_missing_address, n_missing_hours, n_missing_ages)
))
```

## Facility Types

The raw `type` field concatenates program type with offer tags (e.g., `"Child
Care Center Program Type Offers: Faith Based"`). Values are reported as-is
from the scraper; cleaning happens in the standardize stage.

```{r facility-types}
kable(
  raw |> count(type, sort = TRUE),
  caption = "Facility types as reported by CCAV (raw)"
)
```

## Summary

```{r summary}
kable(data.frame(
  Item = c(
    "Source file",
    "Ingest date",
    "Total facilities",
    "Columns"
  ),
  Value = c(
    basename(ingest_path),
    ingest_date,
    as.character(n_total),
    paste(names(raw), collapse = ", ")
  )
))
```

## Data Quality Audit

### Known Issues

```{r audit-known-issues}
cat(sprintf(
  "Address prefix contamination: %d facilities have an address that does not start with a digit.\n",
  n_address_prefix
))
cat("These are cleaned in ccav_standardize.qmd by stripping the director name prefix.\n\n")
cat("Subsidy data absent — CCAV does not report subsidy acceptance. Sourced entirely from DSS.\n")
```

### Missing Fields Summary

```{r audit-missing}
missing_df <- raw |>
  summarise(across(everything(), \(x) sum(is.na(x) | as.character(x) == ""))) |>
  tidyr::pivot_longer(everything(), names_to = "field", values_to = "n_missing") |>
  filter(n_missing > 0) |>
  arrange(desc(n_missing))

if (nrow(missing_df) == 0) {
  cat("No missing values detected.\n")
} else {
  kable(missing_df, caption = "Fields with missing or blank values")
}
```

### Random Spot Check

```{r audit-spot-check}
set.seed(as.integer(Sys.Date()))
kable(
  slice_sample(raw, n = min(5, nrow(raw))) |>
    select(name, licenseType, address, hours, ages),
  caption = "Random date-seeded spot check"
)
```
