---
title: "Hampton Childcare — DSS Ingest"
date: "`r Sys.Date()`"
format: gfm
---

## Overview

Collects all childcare facility records for the City of Hampton (FIPS 650) from
the Virginia Department of Social Services (DSS) facility search at
`https://www.dss.virginia.gov/facility/search/cc2.cgi`. This is a server-side
CGI form — no browser or CAPTCHA required. The DSS list is the authoritative
source for licensure status and subsidy provider designation in Virginia.

**Fields collected:** name, address, phone, facility_type, license_type,
license_id, expiration_date, administrator, hours, capacity, ages, subsidy
provider status, inspector.

**Output:** `Data/ingest/DSS_ingest_YYYY-MM-DD.csv`

| Limitation | Detail |
|---|---|
| SSL verification disabled | Virginia's state server uses a cert chain not trusted by R's default CA bundle. Disabled for this URL only — see note in setup. |
| ~30-second runtime | One GET per facility with a 0.5 s polite delay. |
| Subsidy filter omitted | `search_modifiers_mod_cde=1` intentionally excluded — returns all facilities, not just current subsidy providers. |
| Unlicensed tab | Included; entries may be transitional or out-of-date. |

---

```{r setup, include=FALSE}
library(here)
library(httr)
library(rvest)
library(dplyr)
library(stringr)
library(readr)
library(knitr)

source(here::here("config.R"))
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

DSS_BASE     <- "https://www.dss.virginia.gov/facility/search/cc2.cgi"
HAMPTON_FIPS <- "650"
today        <- format(Sys.Date(), "%Y-%m-%d")
out_file     <- here("Data", "01_ingest", paste0("DSS_ingest_", today, ".csv"))

# Virginia's state server uses a cert chain not in R's default CA bundle.
# SSL verification is disabled for this known government URL only.
# Do not copy this pattern for other URLs.
ssl_no_verify <- config(ssl_verifypeer = 0L, ssl_verifyhost = 0L)

# If a DSS ingest file already exists (e.g. copied from seed), skip scraping
# and report from the existing file. The most recent file is used.
existing_dss <- tail(sort(Sys.glob(here("Data", "01_ingest", "DSS_ingest_*.csv"))), 1)
use_existing <- length(existing_dss) > 0
if (use_existing) {
  message("Existing DSS ingest file found — skipping scrape: ", basename(existing_dss))
  results <- read_csv(existing_dss, show_col_types = FALSE)
}
```

## Helper Functions

```{r helpers, include=FALSE}
# Parse facility links from one result tab (Licensed or Unlicensed)
parse_tab <- function(html, tab_id) {
  tab_div <- html |> html_element(paste0("#", tab_id))
  if (is.na(tab_div)) return(tibble())
  links <- tab_div |> html_elements("a[href*='rm=Details']")
  tibble(
    name = html_text(links, trim = TRUE),
    href = html_attr(links, "href"),
    tab  = tab_id
  ) |> filter(nchar(name) > 0)
}

# Parse a single facility detail page into a one-row tibble.
# The detail data sits in a two-column <table class="cc_search">; each <tr>
# has a label cell and a value cell. Address and phone live in a float:left
# header block where name and street merge after str_squish — street is
# extracted from the first digit onward.
parse_detail <- function(facility_id, name, tab, detail_url) {
  resp <- tryCatch(
    with_config(ssl_no_verify, GET(detail_url)),
    error = function(e) { message("  ERROR fetching ", name, ": ", e$message); NULL }
  )
  if (is.null(resp) || http_error(resp)) {
    message("  SKIP (HTTP error): ", name)
    return(tibble(facility_id = facility_id, name = name, tab = tab, parse_error = TRUE))
  }

  pg <- read_html(content(resp, as = "raw"), encoding = "latin1")

  detail_table <- pg |>
    html_elements("table.cc_search") |>
    (\(tables) {
      keep <- sapply(tables, \(t) str_detect(html_text(t), "Facility Type"))
      if (!any(keep)) return(tables[[1]])
      tables[[which(keep)[1]]]
    })()

  rows <- detail_table |> html_elements("tr")
  kv <- Filter(Negate(is.null), lapply(rows, function(row) {
    cells <- row |> html_elements("td")
    if (length(cells) < 2) return(NULL)
    list(
      label = str_squish(html_text(cells[[1]])),
      value = str_squish(html_text(cells[[2]]))
    )
  }))

  get_val <- function(pattern) {
    matches <- sapply(kv, \(x) str_detect(x$label, regex(pattern, ignore_case = TRUE)))
    if (!any(matches)) return(NA_character_)
    kv[[which(matches)[1]]]$value
  }

  header_trs <- pg |>
    html_elements("table[border='0'] tr td[colspan='2']") |>
    html_text(trim = TRUE) |>
    str_squish()

  street_raw <- header_trs[str_detect(header_trs, "\\d")][1]
  street     <- str_extract(street_raw, "\\d{2,}\\s+[A-Za-z].*")
  city_zip   <- header_trs[str_detect(header_trs, "VA\\s+\\d{5}")][1]
  address <- if_else(
    !is.na(street) & !is.na(city_zip),
    str_squish(paste(street, city_zip, sep = ", ")),
    NA_character_
  )
  phone <- header_trs[str_detect(header_trs, "\\(\\d{3}\\)")][1] |>
    str_extract("\\(\\d{3}\\)\\s*\\d{3}-\\d{4}")

  tibble(
    facility_id,
    name, tab, address, phone,
    facility_type    = get_val("Facility Type"),
    license_type     = get_val("License Type"),
    license_id       = get_val("License/Facility ID"),
    expiration_date  = get_val("Expiration Date"),
    administrator    = get_val("Administrator"),
    hours            = get_val("Business Hours"),
    capacity         = as.integer(str_extract(get_val("Capacity"), "\\d+")),
    ages_raw         = get_val("Ages"),
    subsidy_provider = get_val("Current Subsidy Provider"),
    inspector        = get_val("Inspector"),
    parse_error      = FALSE
  )
}
```

## Fetch Facility List

POST the search form to retrieve all Hampton facilities across the Licensed and
Unlicensed tabs.

```{r fetch-list}
if (!use_existing) {
search_response <- with_config(ssl_no_verify, {
  POST(
    DSS_BASE,
    body   = list(
      rm                   = "Search",
      search_keywords_name = "",
      search_exact_fips    = HAMPTON_FIPS
    ),
    encode = "form"
  )
})
if (http_error(search_response)) {
  stop("Search POST failed: HTTP ", status_code(search_response))
}

search_html <- read_html(content(search_response, as = "raw"), encoding = "latin1")

facility_list <- bind_rows(
  parse_tab(search_html, "Licensed"),
  parse_tab(search_html, "Unlicensed")
) |>
  mutate(
    facility_id = str_match(href, "ID=(\\d+)")[, 2],
    detail_url  = paste0("https://www.dss.virginia.gov", href)
  ) |>
  filter(!is.na(facility_id))

cat(sprintf(
  "Found %d facilities: %d licensed, %d unlicensed\n",
  nrow(facility_list),
  sum(facility_list$tab == "Licensed"),
  sum(facility_list$tab == "Unlicensed")
))
} # end !use_existing
```

## Fetch Detail Pages

One GET request per facility with a 0.5 s delay to avoid overloading the state server.

```{r fetch-details}
if (!use_existing) {
all_records <- vector("list", nrow(facility_list))
for (i in seq_len(nrow(facility_list))) {
  row <- facility_list[i, ]
  message(sprintf("  [%d/%d] %s", i, nrow(facility_list), row$name))
  all_records[[i]] <- parse_detail(row$facility_id, row$name, row$tab, row$detail_url)
  Sys.sleep(0.5)
}
results <- bind_rows(all_records)
} # end !use_existing
```

## Write Output

```{r write-output}
if (!use_existing) {
  dir.create(here("Data", "01_ingest"), showWarnings = FALSE, recursive = TRUE)
  write_csv(results, out_file)
  prune_dated(here("Data", "01_ingest"), "^DSS_ingest_[0-9]{4}-[0-9]{2}-[0-9]{2}\\.csv$")
  cat(sprintf("Wrote %d rows → %s\n", nrow(results), basename(out_file)))
} else {
  cat(sprintf("Reporting from existing file: %s (%d rows)\n", basename(existing_dss), nrow(results)))
}
```

## Main Findings

```{r main-findings}
n_errors  <- sum(results$parse_error, na.rm = TRUE)
n_subsidy <- sum(str_to_lower(results$subsidy_provider) == "yes", na.rm = TRUE)

kable(data.frame(
  Metric = c(
    "Total facilities", "Licensed", "Unlicensed",
    "Current subsidy providers", "Parse errors"
  ),
  Count = c(
    nrow(results),
    sum(results$tab == "Licensed"),
    sum(results$tab == "Unlicensed"),
    n_subsidy,
    n_errors
  )
))
```

## Summary

```{r summary-table}
kable(
  results |>
    count(tab, subsidy_provider) |>
    arrange(tab, subsidy_provider),
  caption = "Facilities by tab and subsidy status"
)
```

## Data Quality Audit

### Parse Errors

```{r audit-parse-errors}
if (n_errors > 0) {
  cat(sprintf("WARNING: %d facilities had parse errors:\n", n_errors))
  for (nm in results$name[results$parse_error == TRUE]) cat(" -", nm, "\n")
} else {
  cat("No parse errors.\n")
}
```

### Missing Fields

```{r audit-missing}
kable(data.frame(
  Field = c("address", "phone", "hours", "capacity", "ages_raw", "subsidy_provider"),
  NA_count = c(
    sum(is.na(results$address)),
    sum(is.na(results$phone)),
    sum(is.na(results$hours)),
    sum(is.na(results$capacity)),
    sum(is.na(results$ages_raw)),
    sum(is.na(results$subsidy_provider))
  )
))
```

### Suspicious Entries

```{r audit-flags}
# Flag facilities where the address does not start with a digit —
# may indicate a name-prefix contamination similar to CCAV
suspicious <- results |>
  filter(!is.na(address), !str_detect(address, "^\\d"))
if (nrow(suspicious) > 0) {
  cat(sprintf("WARNING: %d facilities with non-numeric address start:\n", nrow(suspicious)))
  kable(suspicious |> select(name, address))
} else {
  cat("No suspicious address entries.\n")
}
```

### Random Spot Check

```{r audit-spot-check}
set.seed(as.integer(Sys.Date()))
kable(
  slice_sample(results, n = min(5, nrow(results))) |>
    select(name, tab, address, hours, ages_raw, subsidy_provider),
  caption = "Random date-seeded spot check"
)
```
